---
title: "FRAME2: Block-Group-Level Sampling Frame"
subtitle: "Detroit Transportation Survey Sample Design"
author: "Team Effect Size (Kevin Linares, Tavril Saint Jean)"
date: today
format:
  pdf:
    keep-tex: false
    geometry: margin=0.6in
    fontsize: 10pt
    classoption: landscape
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{array}
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

pacman::p_load(tidyverse, knitr, kableExtra, here, scales)
options(scipen = 999)
here::i_am("inputs/Detroit_MI.xlsx")

ssu_frame <- read_csv(
  here("outputs", "FRAME_BG_EffectSize.csv"),
  col_types = cols(Tract = col_character(),
                   BlockGroup = col_character()),
  show_col_types = FALSE
)
```

This file documents the cleaned block-group-level sampling frame used for the Detroit Transportation Survey. Each row represents one secondary sampling unit (SSU), which is a census block group. For each sampled tract, one block group was selected via systematic probability-proportional-to-size with the BG-level composite measure of size `MOS_bg` driving the conditional selection probability.

The frame contains `r nrow(ssu_frame)` block groups nested within `r length(unique(ssu_frame$Tract))` tracts. The `Tract` column identifies the parent PSU for each BG. Population counts come from the 2020 Decennial Census Demographic and Housing Characteristics Summary File, aggregated from the P12 sex-by-age table into the four survey age domains.

\vspace{0.5em}

```{r}
#| label: tbl-bg-frame

ssu_frame |>
  arrange(Tract, BlockGroup) |>
  select(BlockGroup, Tract, TotHH, TotAdult,
         Age18to34, Age35to49, Age50to64, Age65plus, MOS_bg) |>
  mutate(
    TotHH      = format(TotHH,      big.mark = ",", trim = TRUE),
    TotAdult   = format(TotAdult,   big.mark = ",", trim = TRUE),
    Age18to34  = format(Age18to34,  big.mark = ",", trim = TRUE),
    Age35to49  = format(Age35to49,  big.mark = ",", trim = TRUE),
    Age50to64  = format(Age50to64,  big.mark = ",", trim = TRUE),
    Age65plus  = format(Age65plus,  big.mark = ",", trim = TRUE),
    MOS_bg     = sprintf("%.4f", MOS_bg)
  ) |>
  kable(
    booktabs   = TRUE,
    longtable  = TRUE,
    align      = c("l", "l", rep("r", 7)),
    col.names  = c("BlockGroup FIPS", "Tract FIPS", "TotHH", "TotAdult",
                   "Age 18-34", "Age 35-49", "Age 50-64", "Age 65+", "MOS_bg")
  ) |>
  kable_styling(
    latex_options = c("repeat_header", "striped"),
    font_size     = 10
  )
```
