---
title: "FRAME1: Tract-Level Sampling Frame"
subtitle: "Detroit Transportation Survey Sample Design"
author: "Team Effect Size (Kevin Linares, Tavril Saint Jean)"
date: today
format:
  pdf:
    keep-tex: false
    geometry: margin=0.6in
    fontsize: 10pt
    classoption: landscape
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{array}
execute:
  echo: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

pacman::p_load(tidyverse, knitr, kableExtra, here, scales)
options(scipen = 999)
here::i_am("inputs/Detroit_MI.xlsx")

psu_frame <- read_csv(
  here("outputs", "FRAME_PSU_EffectSize.csv"),
  col_types = cols(Tract = col_character()),
  show_col_types = FALSE
)
```

This file documents the cleaned and Kish-linked tract-level sampling frame used for the Detroit Transportation Survey. Each row represents one primary sampling unit (PSU). The `n_tracts` column indicates how many original Census tracts were combined into each receiving unit (1 means not combined, greater than 1 means the receiving tract absorbed one or more undersized neighbors). The composite measure of size (MOS) and PSU inclusion probability $\pi_i$ were used to select 40 tracts via systematic probability-proportional-to-size sampling with `set.seed(-77)`.

The frame contains `r nrow(psu_frame)` PSUs covering `r format(sum(psu_frame$TotAdult), big.mark = ",")` adults aged 18 and older. Source data are from the 2020 Decennial Census Demographic and Housing Characteristics Summary File, retrieved through the U.S. Census Bureau API.

\vspace{0.5em}

```{r}
#| label: tbl-psu-frame

psu_frame |>
  arrange(Tract) |>
  mutate(
    TotHH      = format(TotHH,      big.mark = ",", trim = TRUE),
    TotAdult   = format(TotAdult,   big.mark = ",", trim = TRUE),
    Age18to34  = format(Age18to34,  big.mark = ",", trim = TRUE),
    Age35to49  = format(Age35to49,  big.mark = ",", trim = TRUE),
    Age50to64  = format(Age50to64,  big.mark = ",", trim = TRUE),
    Age65plus  = format(Age65plus,  big.mark = ",", trim = TRUE),
    MOS  = sprintf("%.4f", MOS),
    pi_i = sprintf("%.6f", pi_i)
  ) |>
  kable(
    booktabs   = TRUE,
    longtable  = TRUE,
    align      = c("l", "c", rep("r", 8)),
    col.names  = c("Tract FIPS", "n_tracts", "TotHH", "TotAdult",
                   "Age 18-34", "Age 35-49", "Age 50-64", "Age 65+",
                   "MOS", "pi_i")
  ) |>
  kable_styling(
    latex_options = c("repeat_header", "striped"),
    font_size     = 10
  )
```
