---
title: "Detroit Transportation Survey: Multi-Stage Area Probability Sample Design"
subtitle: "A Technical Sampling Report"
author:
  - name: "Group Name"
    affiliations:
      - name: "University of Michigan"
date: today
format:
  jasa-pdf:
    keep-tex: true
    toc: true
    toc-depth: 3
    number-sections: true
    cite-method: natbib
  pdf:
    documentclass: article
    geometry:
      - margin=1in
    fontsize: 11pt
    linestretch: 1.0
    toc: true
    toc-depth: 3
    number-sections: true
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{booktabs}
        \usepackage{longtable}
        \usepackage{float}
        \usepackage{amsmath}
        \floatplacement{table}{H}
        \floatplacement{figure}{H}
bibliography: references.bib
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false

# ──────────────────────────────────────────────────────
# Load Required Packages
# ──────────────────────────────────────────────────────
library(tidyverse)      # Data wrangling and visualization
library(readxl)         # Read Excel files
library(sampling)       # Survey sample selection (UPsystematic, etc.)
library(survey)         # Survey analysis
library(srvyr)          # Tidyverse-friendly survey analysis
library(PracTools)      # Practical tools for survey design
library(sf)             # Spatial data handling
library(tigris)         # Census TIGER shapefiles
library(tidycensus)     # Census data API
library(viridis)        # Color-blind friendly palettes
library(patchwork)      # Combine ggplots
library(knitr)          # Table formatting
library(kableExtra)     # Enhanced tables
library(scales)         # Number formatting

# ──────────────────────────────────────────────────────
# API Keys
# Read stored API keys from local CSV
# Adjust path as needed for your machine
# ──────────────────────────────────────────────────────
api_keys <- read_csv("D:/repos/api-keys.csv", show_col_types = FALSE)

census_key <- api_keys |>
  filter(key_id == "Census_key") |>
  pull(key)

stadia_key <- api_keys |>
  filter(key_id == "stadiamaps_key") |>
  pull(key)

# Register the Census API key (install = FALSE to avoid writing to .Renviron)
census_api_key(census_key, install = FALSE)

# Register Stadia Maps key for ggmap
# ggmap::register_stadiamaps(stadia_key)

# ──────────────────────────────────────────────────────
# Tigris Options
# ──────────────────────────────────────────────────────
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

# ──────────────────────────────────────────────────────
# Global knitr options
# ──────────────────────────────────────────────────────
knitr::opts_chunk$set(
  fig.align = "center",
  fig.width = 6.5,
  fig.height = 4.5,
  out.width = "100%"
)
```

\newpage

# Introduction {#sec-intro}

This report describes the design of a multi-stage area probability sample for
a face-to-face survey of Detroit, Michigan residents. The survey examines
transportation security and transportation use, with key outcome variables
including the proportion of residents in households without a vehicle, the
proportion whose main transportation is public transit, the proportion who
skip destinations due to transportation problems, and the proportion who
sought transportation help from friends, family, or neighbors.

The sample uses a three-stage design:

1. **Census tracts** as primary sampling units (PSUs),
2. **Block groups (BGs)** as secondary sampling units (SSUs), and
3. **Persons** as elements.

The design goals are to yield 1,000 total respondents distributed equally
across four age domains (18–34, 35–49, 50–64, 65+), while achieving a
**self-weighting sample within each age group** and **equal interviewer
workload in each sample PSU**. These dual objectives are accomplished
through the **composite measure of size (MOS)** technique described in
Valliant, Dever, and Kreuter (VDK, Chapter 10, Section 10.5).

The report is structured for a dual audience: **managers** seeking a
high-level understanding of the sample design and its rationale, and
**survey statisticians** requiring the mathematical details, selection
probabilities, and variance estimation guidance.

@sec-design presents the sample design, including frame construction and
composite MOS calculations. @sec-selection covers sample selection procedures.
@sec-precision addresses anticipated precision, @sec-weighting describes the
weighting plan, and @sec-variance explains variance estimation. Maps appear
in @sec-maps, and the Appendix provides codebooks and sample listings.


# Sample Design {#sec-design}

## Design Goals

The sample must satisfy four requirements simultaneously:

1. **Total sample size**: 1,000 respondents split equally (250 per group)
   across four age domains.
2. **Self-weighting within domains**: Every person in a given age group
   has the same overall selection probability, i.e., $\pi_k(d) = f_d$ for
   domain $d$.
3. **Equal workload per PSU**: The total expected number of selected persons
   is the same across all 40 sample tracts.
4. **Multi-stage selection**: 40 tracts (PSUs), 1 BG per tract (SSU), and
   persons selected within BGs at domain-specific rates.

## Target Population

The target population consists of all non-institutionalized persons aged 18
and older residing in the City of Detroit, Michigan. According to the 2020
Decennial Census Demographic and Housing Characteristics (DHC) Summary File,
Detroit contains approximately 480,131 adults (18+) across 270 eligible
census tracts.

## Area Sampling Frame {#sec-frame}

### Data Sources

The sampling frame is constructed from two files:

- **Census Tract FIPS Code Detroit MI.csv**: A list of 280 FIPS codes
  identifying Census tracts within the City of Detroit, obtained from the
  City of Detroit Open Data Portal.
- **Detroit MI.xlsx**: Census tract- and block group-level aggregate data from
  the 2020 Decennial Census DHC Summary File, downloaded via the
  `tidycensus` R package for Wayne County, MI and merged with the Detroit
  tract FIPS codes. Includes housing unit counts (`H1_001N`), total person
  counts, and detailed person counts by Sex $\times$ Age (`P12` table).

### Frame Construction

```{r load-data}
#| label: load-data
#| echo: true

# ──────────────────────────────────────────────────────
# 1. Load the Census tract FIPS codes for Detroit
# ──────────────────────────────────────────────────────
tract_fips <- read_csv(
  "Census Tract FIPS Code Detroit MI.csv",
  col_names = "GEOID",
  col_types = cols(GEOID = col_character()),
  skip = 1                    # skip header if present
)

# ──────────────────────────────────────────────────────
# 2. Load tract-level and block-group-level data
# ──────────────────────────────────────────────────────
tract_raw <- read_excel("Detroit MI.xlsx", sheet = "Tract") |>
  mutate(Tract = as.character(Tract))

bg_raw <- read_excel("Detroit MI.xlsx", sheet = "BlockGroup") |>
  mutate(BlockGroup = as.character(BlockGroup))
```

```{r create-domains}
#| label: create-domains
#| echo: true

# ──────────────────────────────────────────────────────
# Helper function: compute age domain counts from
# the detailed Sex x Age columns
# ──────────────────────────────────────────────────────
compute_age_domains <- function(df) {
  df |>
    mutate(
      # Domain 1: Ages 18-34
      Age18_34 = Male18to19Yrs + Male20Yrs + Male21Yrs +
        Male22to24Yrs + Male25to29Yrs + Male30to34Yrs +
        Female18to19Yrs + Female20Yrs + Female21Yrs +
        Female22to24Yrs + Female25to29Yrs + Female30to34Yrs,
      # Domain 2: Ages 35-49
      Age35_49 = Male35to39Yrs + Male40to44Yrs + Male45to49Yrs +
        Female35to39Yrs + Female40to44Yrs + Female45to49Yrs,
      # Domain 3: Ages 50-64
      Age50_64 = Male50to54Yrs + Male55to59Yrs + Male60to61Yrs +
        Male62to64Yrs +
        Female50to54Yrs + Female55to59Yrs + Female60to61Yrs +
        Female62to64Yrs,
      # Domain 4: Ages 65+
      Age65plus = Male65to66Yrs + Male67to69Yrs + Male70to74Yrs +
        Male75to79Yrs + Male80to84Yrs + MaleGE85Yrs +
        Female65to66Yrs + Female67to69Yrs + Female70to74Yrs +
        Female75to79Yrs + Female80to84Yrs + FemaleGE85Yrs,
      # Total eligible adult population (18+)
      AdultPop = Age18_34 + Age35_49 + Age50_64 + Age65plus
    )
}

# ──────────────────────────────────────────────────────
# 3. Apply age domain computation to tracts and BGs
#    De-duplicate tracts (some appear multiple times in
#    the raw file due to BG-level rows for tract 9855)
# ──────────────────────────────────────────────────────
tract_data <- tract_raw |>
  distinct(Tract, .keep_all = TRUE) |>
  compute_age_domains()

bg_data <- bg_raw |>
  compute_age_domains() |>
  mutate(TractID = str_sub(BlockGroup, 1, 11))
```

### Frame Cleaning and Quality Control

Following VDK Chapter 11 (Section 11.1), several quality control checks
are necessary before sample selection:

1. **Remove ineligible tracts**: Tracts with zero adult population are
   institutional or non-residential areas (e.g., industrial zones, water
   features) and are ineligible for a household survey.
2. **Combine undersized units**: Small tracts and BGs that cannot support
   the required workload must be combined with geographic neighbors.
3. **Handle single-BG tracts**: Tracts with only one BG make the SSU
   selection deterministic ($\pi_{j|i} = 1$). These are noted but can
   remain in the frame since the design calls for $\bar{n} = 1$ BG per tract.

```{r frame-qc}
#| label: frame-qc
#| echo: true

# ──────────────────────────────────────────────────────
# 4. Identify and remove zero-adult-population tracts
# ──────────────────────────────────────────────────────
zero_adult_tracts <- tract_data |>
  filter(AdultPop == 0)

cat("Tracts with zero adult population (to be dropped):\n")
zero_adult_tracts |>
  select(Tract, NAME, TotPerson, TotHH, AdultPop) |>
  print(n = Inf)

# Remove zero-adult tracts from frame
tract_frame <- tract_data |>
  filter(AdultPop > 0)

cat("\nEligible tracts remaining:", nrow(tract_frame), "\n")
cat("Total adult population:", format(sum(tract_frame$AdultPop), big.mark = ","), "\n")

# ──────────────────────────────────────────────────────
# 5. Identify very small tracts (candidates for combining)
#    VDK Ch. 11: combine if tract cannot support the
#    workload or would have an extreme selection probability
# ──────────────────────────────────────────────────────
small_tract_threshold <- 50   # tracts with < 50 adults

small_tracts <- tract_frame |>
  filter(AdultPop < small_tract_threshold) |>
  arrange(AdultPop) |>
  select(Tract, NAME, TotHH, AdultPop, Age18_34, Age35_49, Age50_64, Age65plus)

cat("\nSmall tracts (adult pop <", small_tract_threshold, "):\n")
small_tracts |> print(n = Inf)
```

```{r combine-small-tracts}
#| label: combine-small-tracts
#| echo: true

# ──────────────────────────────────────────────────────
# 6. Combine very small tracts with geographic neighbors
#    Following VDK Ch. 11 (Sec. 11.1), small tracts are
#    combined with adjacent tracts to ensure adequate
#    workload and prevent extreme weights.
#
#    Strategy: sort tracts numerically by FIPS code
#    (geographic proxy) and merge small tracts with the
#    next tract on the list.
# ──────────────────────────────────────────────────────

# Define which tracts need combining (adult pop < 50)
tracts_to_combine <- small_tracts$Tract

# Create a mapping: small tract -> neighboring tract
# We sort all tracts and find the nearest neighbor by FIPS code
all_tracts_sorted <- tract_frame |>
  arrange(Tract) |>
  pull(Tract)

# For each small tract, find its position and assign to
# the next non-small tract on the sorted list
create_combine_map <- function(small_tracts_vec, all_tracts_sorted) {
  combine_map <- tibble(
    original = character(),
    combined_into = character()
  )

  for (st in small_tracts_vec) {
    pos <- which(all_tracts_sorted == st)
    # Look forward for a non-small neighbor
    candidates <- all_tracts_sorted[all_tracts_sorted != st]
    candidates <- candidates[!candidates %in% small_tracts_vec]

    if (pos < length(all_tracts_sorted)) {
      # Find closest non-small tract after current position
      forward <- candidates[candidates > st]
      backward <- candidates[candidates < st]
      if (length(forward) > 0) {
        neighbor <- forward[1]
      } else if (length(backward) > 0) {
        neighbor <- tail(backward, 1)
      } else {
        neighbor <- NA_character_
      }
    } else {
      backward <- candidates[candidates < st]
      neighbor <- tail(backward, 1)
    }

    combine_map <- bind_rows(
      combine_map,
      tibble(original = st, combined_into = neighbor)
    )
  }
  combine_map
}

combine_map <- create_combine_map(tracts_to_combine, all_tracts_sorted)

cat("Tract combining assignments:\n")
combine_map |>
  left_join(
    tract_frame |> select(Tract, AdultPop),
    by = c("original" = "Tract")
  ) |>
  rename(original_pop = AdultPop) |>
  print()

# ──────────────────────────────────────────────────────
# 7. Apply combinations: assign a "PSU_ID" to each tract
#    Combined tracts share the same PSU_ID as their
#    neighbor; standalone tracts keep their own FIPS code.
# ──────────────────────────────────────────────────────
tract_frame <- tract_frame |>
  mutate(
    PSU_ID = case_when(
      Tract %in% combine_map$original ~
        combine_map$combined_into[match(Tract, combine_map$original)],
      TRUE ~ Tract
    )
  )

# Aggregate combined tracts to PSU level
psu_frame <- tract_frame |>
  group_by(PSU_ID) |>
  summarise(
    n_tracts_combined = n(),
    TotHH       = sum(TotHH),
    TotPerson   = sum(TotPerson),
    AdultPop    = sum(AdultPop),
    Age18_34    = sum(Age18_34),
    Age35_49    = sum(Age35_49),
    Age50_64    = sum(Age50_64),
    Age65plus   = sum(Age65plus),
    .groups = "drop"
  ) |>
  arrange(PSU_ID)

cat("\nFinal PSU frame:", nrow(psu_frame), "PSUs\n")
cat("Total adult population:", format(sum(psu_frame$AdultPop), big.mark = ","), "\n")
```

Similarly, we clean and organize the block group frame:

```{r bg-frame}
#| label: bg-frame
#| echo: true

# ──────────────────────────────────────────────────────
# 8. Build the block group (SSU) frame
#    - Remove BGs belonging to dropped tracts
#    - Remove zero-adult BGs
#    - Assign PSU_ID from tract-level combinations
# ──────────────────────────────────────────────────────
bg_frame <- bg_data |>
  # Keep only BGs in eligible tracts
  filter(TractID %in% tract_frame$Tract) |>
  # Remove zero-adult BGs
  filter(AdultPop > 0) |>
  # Assign PSU_ID based on tract combining
  mutate(
    PSU_ID = case_when(
      TractID %in% combine_map$original ~
        combine_map$combined_into[match(TractID, combine_map$original)],
      TRUE ~ TractID
    )
  )

cat("Eligible BGs:", nrow(bg_frame), "\n")

# Count BGs per PSU
bgs_per_psu <- bg_frame |>
  count(PSU_ID, name = "n_bgs")

cat("\nBGs per PSU distribution:\n")
summary(bgs_per_psu$n_bgs)

# PSUs with only 1 BG (deterministic SSU selection)
cat("\nPSUs with 1 BG (deterministic SSU):",
    sum(bgs_per_psu$n_bgs == 1), "\n")
```


## Assigning Measures of Size {#sec-mos}

### Design Parameters

The composite MOS method requires specifying the number of persons to be
**selected** (not just respondents) in each domain, accounting for expected
non-response. From VDK Chapter 6, the initial sample size is inflated by the
inverse of the expected response rate:

$$n_d^{\text{select}} = \frac{n_d^{\text{respond}}}{r_d}$$

where $r_d$ is the expected response rate for domain $d$.

```{r design-params}
#| label: design-params
#| echo: true

# ──────────────────────────────────────────────────────
# 9. Define design parameters
# ──────────────────────────────────────────────────────
design_params <- tibble(
  domain     = c("18-34", "35-49", "50-64", "65+"),
  n_respond  = rep(250, 4),
  resp_rate  = c(0.35, 0.45, 0.50, 0.65)
) |>
  mutate(
    # Inflate for non-response
    n_select = n_respond / resp_rate,
    # Population counts from frame
    N_d = c(
      sum(psu_frame$Age18_34),
      sum(psu_frame$Age35_49),
      sum(psu_frame$Age50_64),
      sum(psu_frame$Age65plus)
    ),
    # Domain sampling rates: f_d = n_select / N_d (VDK Eq. 10.1)
    f_d = n_select / N_d,
    # Percent of adult population
    pct_pop = N_d / sum(N_d) * 100
  )

# Total expected selections and workload per PSU
m <- 40     # number of sample PSUs
n_bar <- 1  # BGs per sample PSU
total_select <- sum(design_params$n_select)
q_bar <- total_select / (m * n_bar)   # workload per SSU (VDK notation)

cat("Total selections needed:", round(total_select, 1), "\n")
cat("Workload per PSU (q-bar):", round(q_bar, 1), "\n")

# Display design parameters
design_params |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  kable(
    col.names = c("Domain", "Respondents", "Resp. Rate",
                   "Selections", "Population (N_d)",
                   "Rate (f_d)", "% Pop"),
    caption = "Design parameters by age domain. Selections are inflated for non-response.",
    digits = c(0, 0, 2, 1, 0, 6, 1),
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position"))
```

### Composite Measure of Size

Following VDK Equation 10.1, the composite MOS for PSU $i$ is:

$$S_i = \sum_{d=1}^{D} f_d \cdot N_i(d) \tag{1}$$

where $f_d$ is the domain sampling rate and $N_i(d)$ is the population count
for domain $d$ in PSU $i$. The total MOS across all PSUs equals the total
expected sample size: $S = \sum_{i \in U} S_i = \sum_{d} f_d N_d = n^{\text{select}}$.

Similarly, the composite MOS for BG $j$ in PSU $i$ (VDK Ch. 11) is:

$$S_{ij} = \sum_{d=1}^{D} f_d \cdot Q_{ij}(d)$$

where $Q_{ij}(d)$ is the population in domain $d$ in BG $j$ of tract $i$.

```{r compute-mos}
#| label: compute-mos
#| echo: true

# ──────────────────────────────────────────────────────
# 10. Compute composite MOS for PSUs and BGs
# ──────────────────────────────────────────────────────

# Extract domain sampling rates as a named vector
f_d <- design_params$f_d
names(f_d) <- design_params$domain

# PSU-level composite MOS
psu_frame <- psu_frame |>
  mutate(
    MOS = f_d["18-34"] * Age18_34 +
      f_d["35-49"] * Age35_49 +
      f_d["50-64"] * Age50_64 +
      f_d["65+"]   * Age65plus
  )

S_total <- sum(psu_frame$MOS)
cat("Total PSU MOS (S):", round(S_total, 2), "\n")
cat("Equals total selections:", round(total_select, 2), "\n")

# BG-level composite MOS
bg_frame <- bg_frame |>
  mutate(
    MOS_bg = f_d["18-34"] * Age18_34 +
      f_d["35-49"] * Age35_49 +
      f_d["50-64"] * Age50_64 +
      f_d["65+"]   * Age65plus
  )

# PSU selection probabilities: pi_i = m * S_i / S (VDK Eq. 10.1)
psu_frame <- psu_frame |>
  mutate(pi_i = m * MOS / S_total)

cat("\nPSU selection probability range:",
    round(min(psu_frame$pi_i), 4), "to",
    round(max(psu_frame$pi_i), 4), "\n")
cat("Any pi_i > 1?", any(psu_frame$pi_i > 1), "\n")
```

### Descriptive Statistics on the Frame

```{r frame-descriptives}
#| label: frame-descriptives
#| echo: true

# ──────────────────────────────────────────────────────
# 11. Summary statistics for the PSU frame
# ──────────────────────────────────────────────────────
psu_summary <- psu_frame |>
  summarise(
    across(
      c(TotHH, AdultPop, Age18_34, Age35_49, Age50_64, Age65plus, MOS, pi_i),
      list(
        min    = \(x) min(x),
        median = \(x) median(x),
        mean   = \(x) mean(x),
        max    = \(x) max(x),
        sd     = \(x) sd(x)
      ),
      .names = "{.col}__{.fn}"
    )
  ) |>
  pivot_longer(
    everything(),
    names_to = c("variable", "statistic"),
    names_sep = "__"
  ) |>
  pivot_wider(names_from = statistic, values_from = value)

psu_summary |>
  mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  kable(
    caption = "Descriptive statistics for the PSU (tract) frame.",
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r frame-histograms}
#| label: fig-frame-hist
#| fig-cap: "Distribution of composite MOS and adult population across PSUs."
#| echo: true
#| fig-height: 3.5

# ──────────────────────────────────────────────────────
# 12. Visualize the frame
# ──────────────────────────────────────────────────────
p1 <- ggplot(psu_frame, aes(x = MOS)) +
  geom_histogram(bins = 30, fill = viridis(1, begin = 0.4), color = "white") +
  labs(x = "Composite MOS", y = "Count",
       title = "Composite MOS Distribution") +
  theme_minimal(base_size = 10)

p2 <- ggplot(psu_frame, aes(x = AdultPop)) +
  geom_histogram(bins = 30, fill = viridis(1, begin = 0.7), color = "white") +
  labs(x = "Adult Population (18+)", y = "Count",
       title = "Adult Population Distribution") +
  theme_minimal(base_size = 10)

p1 + p2
```

## Discussion: Could a Uniform Sampling Rate Be Used? {#sec-uniform}

A natural question is whether a single sampling rate $f$ applied to all adults,
regardless of age, could meet the domain sample size targets. If $f$ is
the same for all domains, then $f = n^{\text{select}} / N_{\text{adult}}$.

```{r uniform-rate}
#| label: uniform-rate
#| echo: true

# ──────────────────────────────────────────────────────
# 13. Evaluate a uniform-rate design
# ──────────────────────────────────────────────────────
N_adult <- sum(psu_frame$AdultPop)
f_uniform <- total_select / N_adult

uniform_eval <- design_params |>
  mutate(
    expected_select_uniform = f_uniform * N_d,
    expected_respond_uniform = expected_select_uniform * resp_rate,
    shortfall = n_respond - expected_respond_uniform,
    pct_shortfall = shortfall / n_respond * 100
  )

uniform_eval |>
  select(domain, N_d, expected_select_uniform,
         expected_respond_uniform, shortfall, pct_shortfall) |>
  mutate(across(where(is.numeric), \(x) round(x, 1))) |>
  kable(
    col.names = c("Domain", "Population", "Exp. Selections",
                   "Exp. Respondents", "Shortfall", "% Shortfall"),
    caption = "Expected outcomes under a uniform sampling rate design.",
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position"))
```

A uniform rate of $f =$ `r round(f_uniform, 6)` would yield approximately
equal selections per person regardless of age. However, because response
rates differ substantially across age groups (from 0.35 for ages 18–34 to
0.65 for ages 65+), the expected number of **respondents** would deviate
from the target of 250 per domain. Specifically, the 35–49 age group would
fall approximately 8.5% short of the target. The composite MOS approach
avoids this problem by building the differential response rates into the
design from the start, ensuring each domain receives the correct number
of selections to meet the respondent targets after non-response.

**Advantages of uniform rate**: simpler design, equal weights for all
persons, no domain identification required at selection.

**Disadvantages**: cannot meet domain-specific respondent targets when
response rates vary; precision in low-response domains is compromised.


# Sample Selection {#sec-selection}

## Method of Selection

The assignment specifies **systematic sampling with probabilities
proportional to size (PPS)**. We use the `UPsystematic` function from
the R `sampling` package, which implements Madow's systematic PPS method.
This is applied at both stages:

1. **Stage 1 (PSUs)**: Select $m = 40$ tracts with probabilities proportional
   to the composite MOS, $\pi_i = m \cdot S_i / S$.
2. **Stage 2 (SSUs)**: Within each sample tract, select $\bar{n} = 1$ BG
   with probability proportional to its BG-level composite MOS,
   $\pi_{j|i} = \bar{n} \cdot S_{ij} / S_i$.

The frame is sorted by tract FIPS code prior to selection, providing
implicit geographic stratification (VDK Section 3.3). The random seed
`set.seed(-77)` is used for reproducibility as required.

**Note on systematic PPS**: Systematic PPS does not permit computation of
joint selection probabilities ($\pi_{ij}$), which precludes certain
unbiased variance estimators. In practice, the selection intervals are
collapsed into analytic strata and a with-replacement variance estimator
is applied (VDK Section 3.3; see @sec-variance).

## Stage 1: PSU Selection

```{r psu-selection}
#| label: psu-selection
#| echo: true

# ──────────────────────────────────────────────────────
# 14. Select PSU sample via systematic PPS
#     (VDK Ch. 10; R sampling package)
# ──────────────────────────────────────────────────────

# Sort frame by tract FIPS code for implicit stratification
psu_frame <- psu_frame |> arrange(PSU_ID)

# Normalized inclusion probabilities (must sum to m = 40)
psu_pik <- psu_frame$pi_i

# Verify inclusion probabilities sum to m
cat("Sum of pi_i:", round(sum(psu_pik), 4), "(should be", m, ")\n")

# Select PSUs using systematic PPS
set.seed(-77)
psu_selected_idx <- UPsystematic(pik = psu_pik)

# Extract selected PSUs (indicator = 1)
psu_sample <- psu_frame |>
  mutate(selected = psu_selected_idx) |>
  filter(selected == 1)

cat("Number of PSUs selected:", nrow(psu_sample), "\n")
```

```{r psu-sample-table}
#| label: psu-sample-table
#| echo: true

# ──────────────────────────────────────────────────────
# 15. Display selected PSU characteristics
# ──────────────────────────────────────────────────────
psu_sample |>
  select(PSU_ID, AdultPop, Age18_34, Age35_49, Age50_64, Age65plus,
         MOS, pi_i) |>
  mutate(
    base_weight = 1 / pi_i,
    across(c(MOS, pi_i, base_weight), \(x) round(x, 4))
  ) |>
  kable(
    caption = "Selected PSUs with population counts, composite MOS, and selection probabilities.",
    booktabs = TRUE,
    longtable = TRUE
  ) |>
  kable_styling(
    latex_options = c("hold_position", "repeat_header", "scale_down"),
    font_size = 8
  )
```

## Stage 2: SSU (Block Group) Selection

Within each of the 40 sample tracts, we select $\bar{n} = 1$ block group
with probability proportional to the BG-level composite MOS. For tracts
with a single BG, the selection is deterministic ($\pi_{j|i} = 1$).

```{r ssu-selection}
#| label: ssu-selection
#| echo: true

# ──────────────────────────────────────────────────────
# 16. Select SSUs (BGs) within each sample PSU
#     One BG per PSU, PPS to BG-level composite MOS
# ──────────────────────────────────────────────────────

# Get BGs belonging to sample PSUs
bg_in_sample_psus <- bg_frame |>
  filter(PSU_ID %in% psu_sample$PSU_ID)

# For each sample PSU, compute conditional BG selection probability
# pi_{j|i} = n_bar * S_ij / S_i
bg_in_sample_psus <- bg_in_sample_psus |>
  left_join(
    psu_sample |> select(PSU_ID, MOS_psu = MOS),
    by = "PSU_ID"
  ) |>
  mutate(
    pij_given_i = n_bar * MOS_bg / MOS_psu
  )

# Select 1 BG per PSU using systematic PPS within each PSU
set.seed(-77)   # reset seed for SSU selection

select_one_bg <- function(bg_subset) {
  # If only 1 BG, selection is certain

  if (nrow(bg_subset) == 1) {
    return(bg_subset |> mutate(ssu_selected = 1L))
  }

  # Systematic PPS within PSU
  pik_bg <- bg_subset$pij_given_i

  # Ensure probabilities are valid (sum to n_bar = 1)
  pik_bg <- pik_bg / sum(pik_bg) * n_bar

  # Clamp any probabilities > 1 to 1 (certainty selections)
  pik_bg <- pmin(pik_bg, 1)

  selected <- UPsystematic(pik = pik_bg)
  bg_subset |> mutate(ssu_selected = as.integer(selected))
}

# Apply selection within each PSU using map (no for loops)
bg_selected <- bg_in_sample_psus |>
  group_by(PSU_ID) |>
  group_split() |>
  map_dfr(select_one_bg) |>
  filter(ssu_selected == 1)

cat("BGs selected:", nrow(bg_selected), "\n")
```

## Selection Probabilities and Base Weights

The overall selection probability for a person $k$ in domain $d$ within BG $j$
of tract $i$ in a self-weighting design is (VDK Eq. 10.5, Ch. 10):

$$\pi_k(d) = \pi_i \cdot \pi_{j|i} \cdot \pi_{k|ij}(d) = f_d$$

The within-SSU sampling rate for domain $d$ is derived from the self-weighting
requirement (VDK Section 10.5.1):

$$\pi_{k|ij}(d) = \frac{f_d}{\pi_i \cdot \pi_{j|i}} = \frac{\bar{q} \cdot f_d}{S_{ij}}$$

where $\bar{q}$ is the target workload per SSU.

```{r sel-probs}
#| label: sel-probs
#| echo: true

# ──────────────────────────────────────────────────────
# 17. Compute selection probabilities and within-SSU
#     sampling rates for persons in each domain
# ──────────────────────────────────────────────────────

# Join PSU selection probabilities to selected BGs
sample_bgs <- bg_selected |>
  left_join(
    psu_sample |> select(PSU_ID, pi_i),
    by = "PSU_ID"
  ) |>
  mutate(
    # Overall BG selection probability
    pi_ij = pi_i * pij_given_i,
    # Within-BG sampling rates for each domain (VDK Eq. 10.2)
    rate_18_34  = q_bar * f_d["18-34"]  / MOS_bg,
    rate_35_49  = q_bar * f_d["35-49"]  / MOS_bg,
    rate_50_64  = q_bar * f_d["50-64"]  / MOS_bg,
    rate_65plus = q_bar * f_d["65+"]    / MOS_bg,
    # Expected sample sizes per domain in this BG
    exp_n_18_34  = rate_18_34  * Age18_34,
    exp_n_35_49  = rate_35_49  * Age35_49,
    exp_n_50_64  = rate_50_64  * Age50_64,
    exp_n_65plus = rate_65plus * Age65plus,
    # Total expected workload in this BG
    exp_workload = exp_n_18_34 + exp_n_35_49 + exp_n_50_64 + exp_n_65plus
  )

cat("Expected workload per BG (should be ~",
    round(q_bar, 1), "):\n")
summary(sample_bgs$exp_workload)

# Verify self-weighting: overall selection probabilities
cat("\nVerification of self-weighting property:\n")
cat("pi_i * pi_{j|i} * rate_18_34 for first BG:",
    round(sample_bgs$pi_i[1] * sample_bgs$pij_given_i[1] * sample_bgs$rate_18_34[1], 6),
    "  (should be f_d[18-34] =", round(f_d["18-34"], 6), ")\n")
```

```{r qc-checks}
#| label: qc-checks
#| echo: true

# ──────────────────────────────────────────────────────
# 18. Quality control checks (VDK Ch. 11, Sec. 11.1)
#     Verify: q*_ij(d) <= Q_ij(d) for all SSUs and domains
#     i.e., we don't try to sample more than the pop.
# ──────────────────────────────────────────────────────

qc_check <- sample_bgs |>
  mutate(
    check_18_34  = exp_n_18_34  <= Age18_34,
    check_35_49  = exp_n_35_49  <= Age35_49,
    check_50_64  = exp_n_50_64  <= Age50_64,
    check_65plus = exp_n_65plus <= Age65plus,
    check_total  = exp_workload <= AdultPop,
    # Also check within-BG rates are <= 1
    rate_ok_18_34  = rate_18_34  <= 1,
    rate_ok_35_49  = rate_35_49  <= 1,
    rate_ok_50_64  = rate_50_64  <= 1,
    rate_ok_65plus = rate_65plus <= 1
  )

# Summarize violations
violations <- qc_check |>
  summarise(
    across(starts_with("check_"), \(x) sum(!x)),
    across(starts_with("rate_ok_"), \(x) sum(!x))
  )

cat("QC violations (number of BGs violating each constraint):\n")
violations |>
  pivot_longer(everything(), names_to = "check", values_to = "violations") |>
  print(n = Inf)
```

## How Persons Should Be Selected

Within each selected block group, field interviewers will compile a listing
of all households and persons aged 18+. Persons are then sampled at the
domain-specific rates $\pi_{k|ij}(d) = \bar{q} f_d / S_{ij}$. Because the
rates are generally not simple fractions, **Poisson sampling** or
**systematic sampling** at these rates should be used rather than selecting
fixed counts (VDK Section 10.5.1).

The procedure is:

1. List all persons 18+ in the selected BG.
2. Classify each person into one of the four age domains.
3. Within each domain, select persons independently with probability
   $\pi_{k|ij}(d)$ (Poisson sampling) or select a systematic sample at
   the specified rate from a randomized list.


# Anticipated Precision {#sec-precision}

## Precision of Estimated Proportions

For the key survey variable---the proportion of persons whose main means of
transportation is public transit---we assess anticipated precision under the
sample design.

Under a self-weighting design, the variance of an estimated proportion
$\hat{p}$ in a cluster sample can be approximated using the design effect:

$$\text{Var}(\hat{p}) \approx \text{deff} \cdot \frac{p(1-p)}{n}$$

where $n$ is the effective respondent sample size and $\text{deff}$ accounts
for the clustering and unequal weighting.

```{r precision}
#| label: precision
#| echo: true

# ──────────────────────────────────────────────────────
# 19. Anticipated precision for proportion estimates
#     Using a range of assumed proportions and deffs
# ──────────────────────────────────────────────────────

# Assume a range of plausible proportions for public transit use
p_values <- c(0.10, 0.15, 0.20, 0.25, 0.30)

# Design effects: VDK Ch. 9 suggests deff depends on
# intraclass correlation (delta) and cluster size.
# For a 3-stage design with m=40 PSUs and ~25 persons per PSU:
# deff ≈ 1 + (n_bar_cluster - 1) * delta
# With delta ~ 0.02-0.05 for transportation variables and
# ~25 persons per cluster: deff ≈ 1.5 to 2.2
deff_values <- c(1.5, 2.0, 2.5)

# Overall sample size of respondents
n_total <- 1000

precision_table <- expand_grid(
  p = p_values,
  deff = deff_values
) |>
  mutate(
    var_p    = deff * p * (1 - p) / n_total,
    se_p     = sqrt(var_p),
    cv_p     = se_p / p,
    moe_95   = 1.96 * se_p,
    eff_n    = n_total / deff
  )

precision_table |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  select(p, deff, se_p, cv_p, moe_95) |>
  kable(
    col.names = c("p", "deff", "SE(p)", "CV(p)", "95% MoE"),
    caption = "Anticipated precision for estimated proportions of public transit use.",
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position"))
```

For a transit use proportion of approximately 0.20 (plausible for Detroit),
the 95% margin of error ranges from $\pm$0.025 (deff = 1.5) to $\pm$0.031
(deff = 2.5), which would be adequate for most policy analyses.

For **domain-level** estimates (e.g., transit use among ages 18–34 only),
the effective sample size is approximately 250 respondents. The margin of
error approximately doubles compared to the full-sample estimate.

```{r precision-domain}
#| label: precision-domain
#| echo: true

# Domain-level precision (n = 250 per domain)
domain_precision <- tibble(
  domain = c("18-34", "35-49", "50-64", "65+"),
  n_respond = rep(250, 4),
  p = 0.20,
  deff = 2.0
) |>
  mutate(
    se_p = sqrt(deff * p * (1 - p) / n_respond),
    moe_95 = 1.96 * se_p,
    cv = se_p / p
  )

domain_precision |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  kable(
    col.names = c("Domain", "n", "p", "deff", "SE(p)", "95% MoE", "CV"),
    caption = "Anticipated domain-level precision (assuming p = 0.20, deff = 2.0).",
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position"))
```


# Weighting Plan {#sec-weighting}

## Overview for Managers

Survey weights are numbers assigned to each respondent that account for
the probability of being selected into the sample. They allow us to produce
estimates that are representative of the entire Detroit adult population,
not just the 1,000 persons who respond. The weighting process involves
three steps:

1. **Base weight**: The inverse of each person's selection probability.
2. **Non-response adjustment**: Adjusts weights upward within each age group
   to compensate for persons who were selected but did not respond.
3. **Calibration (post-stratification)**: Adjusts weights so that weighted
   totals match known population counts from the Census.

## Technical Details

### Base Weights

Because the design is self-weighting within each age domain, the base
weight for every person in domain $d$ is:

$$w_k^{\text{base}}(d) = \frac{1}{\pi_k(d)} = \frac{1}{f_d} = \frac{N_d}{n_d^{\text{select}}}$$

```{r base-weights}
#| label: base-weights
#| echo: true

# ──────────────────────────────────────────────────────
# 20. Base weights by domain
# ──────────────────────────────────────────────────────
base_weights <- design_params |>
  mutate(
    base_weight = 1 / f_d,
    weighted_pop = n_select * base_weight  # should equal N_d
  ) |>
  select(domain, f_d, base_weight, N_d, weighted_pop)

base_weights |>
  mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  kable(
    col.names = c("Domain", "f_d", "Base Weight", "N_d", "Weighted N"),
    caption = "Base weights by age domain under the self-weighting design.",
    booktabs = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position"))
```

### Non-Response Adjustment

Within each age domain, the non-response adjustment factor is the inverse
of the weighted response rate within the domain. If the response rate is
approximately constant within each domain, the adjustment simplifies to:

$$w_k^{\text{NR}}(d) = w_k^{\text{base}}(d) \cdot \frac{1}{r_d}$$

This ensures that the weighted sum of respondents still estimates the
domain population total $N_d$.

### Calibration

Post-stratification or raking calibration can further align the weighted
sample to known Census totals for age groups, sex, race/ethnicity, and
geography (VDK Chapters 13--14). For example, the weights can be
calibrated so that the weighted totals of adults aged 18--34, 35--49,
50--64, and 65+ match the Census DHC counts exactly.


# Variance Estimation {#sec-variance}

## Overview for Managers

Because the sample is selected in stages (tracts, then block groups, then
persons), the variability in survey estimates comes from multiple sources.
Standard statistical formulas that assume simple random sampling would
underestimate the true uncertainty. Special variance estimation methods are
needed.

## Technical Guidance for Analysts

### Defining the Analysis Structure

Analysts using the `survey` package in R (Lumley, 2004) or similar
software should define the analysis structure as follows:

- **Strata**: For variance estimation with a systematic PPS sample, the
  convention is to form **analytic strata** by pairing adjacent sample PSUs
  in the selection order. This creates approximately $m/2 = 20$ analytic
  strata, each containing 2 PSUs (VDK Section 3.3 and Chapter 15).
- **PSUs (clusters)**: The 40 selected census tracts.
- **Weights**: The final calibrated weights $w_k^{\text{final}}$.

```{r variance-setup}
#| label: variance-setup
#| echo: true

# ──────────────────────────────────────────────────────
# 21. Create analytic strata for variance estimation
#     Pair adjacent sample PSUs in selection order
# ──────────────────────────────────────────────────────

# PSUs are already in selection order (sorted by tract FIPS)
psu_sample_ve <- psu_sample |>
  arrange(PSU_ID) |>
  mutate(
    psu_order = row_number(),
    # Create analytic strata by pairing: PSUs 1-2 = stratum 1, etc.
    analytic_stratum = ceiling(psu_order / 2)
  )

cat("Number of analytic strata:", max(psu_sample_ve$analytic_stratum), "\n")
cat("PSUs per stratum:", table(psu_sample_ve$analytic_stratum), "\n")
```

### R Code Template for Analysts

Analysts should set up the survey design object as follows (once
microdata are collected):

```{r survey-design-template}
#| label: survey-design-template
#| eval: false
#| echo: true

# ──────────────────────────────────────────────────────
# Template: survey design object for analysis
# (to be used after data collection)
# ──────────────────────────────────────────────────────

# Assuming 'survey_data' contains respondent records with:
#   - psu_id: the sample tract (PSU) identifier
#   - analytic_stratum: the variance stratum (paired PSUs)
#   - final_weight: the final calibrated weight

detroit_design <- svydesign(
  id       = ~ psu_id,           # PSU (tract) identifier
  strata   = ~ analytic_stratum, # paired-PSU variance strata
  weights  = ~ final_weight,     # final analysis weights
  data     = survey_data,
  nest     = TRUE                # PSU IDs nest within strata
)

# Example: estimate proportion using public transit
svymean(~ transit_main, design = detroit_design)
svyciprop(~ transit_main, design = detroit_design, method = "logit")
```


# Maps {#sec-maps}

```{r load-shapefiles}
#| label: load-shapefiles
#| echo: true
#| cache: true

# ──────────────────────────────────────────────────────
# 22. Load shapefiles for mapping
# ──────────────────────────────────────────────────────
detroit_tracts_sf <- tracts(state = "MI", county = "Wayne", year = 2020) |>
  right_join(tract_fips, by = "GEOID") |>
  st_transform(crs = 4326)

detroit_bgs_sf <- block_groups(state = "MI", county = "Wayne", year = 2020) |>
  mutate(GEOIDt = str_sub(GEOID, 1, 11)) |>
  right_join(tract_fips, by = c("GEOIDt" = "GEOID")) |>
  select(-GEOIDt) |>
  st_transform(crs = 4326)
```

```{r map-detroit}
#| label: fig-map-detroit
#| fig-cap: "City of Detroit census tracts (solid) and block groups (dotted)."
#| echo: true
#| fig-height: 6

# ──────────────────────────────────────────────────────
# 23. Map of all Detroit tracts and block groups
# ──────────────────────────────────────────────────────
ggplot() +
  geom_sf(data = detroit_bgs_sf, fill = NA,
          linetype = "dotted", linewidth = 0.3, color = "grey50") +
  geom_sf(data = detroit_tracts_sf, fill = NA,
          color = "black", linewidth = 0.5) +
  labs(
    title = "City of Detroit, Michigan",
    subtitle = "Census Tracts (solid) and Block Groups (dotted)",
    caption = "Source: 2020 TIGER/Line Shapefiles via tigris"
  ) +
  theme_void(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

```{r map-sample}
#| label: fig-map-sample
#| fig-cap: "Selected sample tracts (blue fill) and selected block groups (red boundary) within Detroit."
#| echo: true
#| fig-height: 6

# ──────────────────────────────────────────────────────
# 24. Highlight sample PSUs and SSUs on the map
# ──────────────────────────────────────────────────────

# Identify sample tracts in the shapefile
sample_tracts_sf <- detroit_tracts_sf |>
  filter(GEOID %in% psu_sample$PSU_ID)

# Identify selected BGs in the shapefile
sample_bgs_sf <- detroit_bgs_sf |>
  filter(GEOID %in% sample_bgs$BlockGroup)

ggplot() +
  # All tracts (background)
  geom_sf(data = detroit_tracts_sf, fill = NA,
          color = "grey70", linewidth = 0.3) +
  # Selected tracts (blue fill)
  geom_sf(data = sample_tracts_sf,
          fill = viridis(1, begin = 0.4, alpha = 0.3),
          color = viridis(1, begin = 0.4), linewidth = 0.6) +
  # Selected BGs (red boundary)
  geom_sf(data = sample_bgs_sf, fill = NA,
          color = "red", linewidth = 0.8) +
  labs(
    title = "Selected Sample: 40 Tracts and Block Groups",
    subtitle = "Blue = sample tracts; Red outline = selected block groups",
    caption = "Source: 2020 TIGER/Line Shapefiles via tigris"
  ) +
  theme_void(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

```{r map-mos}
#| label: fig-map-mos
#| fig-cap: "Choropleth of composite MOS across Detroit tracts."
#| echo: true
#| fig-height: 6

# ──────────────────────────────────────────────────────
# 25. Choropleth showing composite MOS
# ──────────────────────────────────────────────────────
mos_map_data <- detroit_tracts_sf |>
  left_join(psu_frame |> select(PSU_ID, MOS), by = c("GEOID" = "PSU_ID"))

ggplot() +
  geom_sf(data = mos_map_data, aes(fill = MOS),
          color = "white", linewidth = 0.2) +
  scale_fill_viridis_c(
    name = "Composite\nMOS",
    option = "viridis",
    na.value = "grey90"
  ) +
  labs(
    title = "Composite Measure of Size by Census Tract",
    subtitle = "Higher MOS = higher selection probability",
    caption = "Source: 2020 DHC Summary File; Authors' calculations"
  ) +
  theme_void(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```


# Appendix {#sec-appendix}

## A. Codebook: PSU (Tract) Frame File {#sec-codebook-psu}

```{r codebook-psu}
#| label: codebook-psu
#| echo: false

tribble(
  ~Variable, ~Type, ~Description,
  "PSU_ID",       "Character (11)", "Census tract FIPS code (PSU identifier)",
  "n_tracts_combined", "Integer", "Number of original tracts combined into this PSU",
  "TotHH",        "Integer", "Total housing units from 2020 Census (H1_001N)",
  "TotPerson",    "Integer", "Total person count from 2020 Census",
  "AdultPop",     "Integer", "Total adult (18+) population",
  "Age18_34",     "Integer", "Population aged 18-34 (domain 1)",
  "Age35_49",     "Integer", "Population aged 35-49 (domain 2)",
  "Age50_64",     "Integer", "Population aged 50-64 (domain 3)",
  "Age65plus",    "Integer", "Population aged 65+ (domain 4)",
  "MOS",          "Numeric", "Composite measure of size: sum_d(f_d * N_i(d))",
  "pi_i",         "Numeric", "PSU selection probability: m * S_i / S"
) |>
  kable(caption = "Codebook for the PSU (tract-level) frame file.",
        booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

## B. Codebook: SSU (Block Group) Sample File {#sec-codebook-ssu}

```{r codebook-ssu}
#| label: codebook-ssu
#| echo: false

tribble(
  ~Variable, ~Type, ~Description,
  "BlockGroup",   "Character (12)", "Block group FIPS code (SSU identifier)",
  "PSU_ID",       "Character (11)", "Parent PSU (tract) identifier",
  "TotHH",        "Integer", "Total housing units in the BG",
  "AdultPop",     "Integer", "Total adult (18+) population in the BG",
  "Age18_34",     "Integer", "BG population aged 18-34",
  "Age35_49",     "Integer", "BG population aged 35-49",
  "Age50_64",     "Integer", "BG population aged 50-64",
  "Age65plus",    "Integer", "BG population aged 65+",
  "MOS_bg",       "Numeric", "BG-level composite MOS: sum_d(f_d * Q_ij(d))",
  "pi_i",         "Numeric", "PSU selection probability",
  "pij_given_i",  "Numeric", "Conditional BG selection probability: S_ij / S_i",
  "pi_ij",        "Numeric", "Joint selection probability: pi_i * pi_{j|i}",
  "rate_18_34",   "Numeric", "Within-BG person sampling rate, ages 18-34",
  "rate_35_49",   "Numeric", "Within-BG person sampling rate, ages 35-49",
  "rate_50_64",   "Numeric", "Within-BG person sampling rate, ages 50-64",
  "rate_65plus",  "Numeric", "Within-BG person sampling rate, ages 65+",
  "exp_workload", "Numeric", "Expected total selections in this BG (~q_bar)"
) |>
  kable(caption = "Codebook for the SSU (block group) sample file.",
        booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

## C. Sample PSU and SSU Listing {#sec-sample-listing}

```{r sample-listing}
#| label: sample-listing
#| echo: true

# ──────────────────────────────────────────────────────
# 26. Full listing of sample PSUs and SSUs with
#     selection probabilities and domain sampling rates
# ──────────────────────────────────────────────────────
sample_listing <- sample_bgs |>
  select(
    PSU_ID, BlockGroup, AdultPop,
    Age18_34, Age35_49, Age50_64, Age65plus,
    MOS_bg, pi_i, pij_given_i, pi_ij,
    rate_18_34, rate_35_49, rate_50_64, rate_65plus,
    exp_workload
  ) |>
  arrange(PSU_ID)

sample_listing |>
  mutate(across(where(is.numeric), \(x) round(x, 4))) |>
  kable(
    caption = "Selected sample PSUs and SSUs with selection probabilities and within-BG domain sampling rates.",
    booktabs = TRUE,
    longtable = TRUE
  ) |>
  kable_styling(
    latex_options = c("hold_position", "repeat_header", "scale_down"),
    font_size = 7
  )
```

## D. Exported Frame and Sample Files

```{r export-files}
#| label: export-files
#| echo: true

# ──────────────────────────────────────────────────────
# 27. Export frame and sample files
# ──────────────────────────────────────────────────────

# PSU frame file
write_csv(psu_frame, "FRAME_PSU_GROUPNAME.csv")

# BG frame file (all BGs in eligible tracts)
write_csv(bg_frame, "FRAME_BG_GROUPNAME.csv")

# Sample file (selected BGs with all probabilities)
write_csv(sample_listing, "SAMPLE_GROUPNAME.csv")

cat("Files exported successfully.\n")
```

## E. Disclosure of AI Use

This project utilized **Claude (Anthropic, Opus 4.6 model)** as an AI
assistant for the following purposes:

1. **Code development**: Generating and debugging R/Quarto code for frame
   construction, composite MOS computation, systematic PPS sample selection,
   and map visualization.
2. **Report drafting**: Assisting with structuring the technical report,
   writing narrative explanations for both managers and technical audiences,
   and formatting equations.
3. **Verification**: Cross-checking calculations (domain population totals,
   sampling rates, selection probabilities, workload computations) against
   the VDK textbook formulas.

All prompts to the AI assistant are documented in the project files.
The AI's outputs were reviewed, verified, and modified by team members
to ensure accuracy and alignment with course materials.

\newpage

# References {.unnumbered}

::: {#refs}
Valliant, R., Dever, J. A., & Kreuter, F. (2018). *Practical Tools for
Designing and Weighting Survey Samples* (2nd ed.). Springer.

Lumley, T. (2004). Analysis of complex survey samples. *Journal of
Statistical Software*, 9(8), 1–19.

Sampford, M. R. (1967). On sampling without replacement with unequal
probabilities of selection. *Biometrika*, 54(3-4), 499–513.

Tillé, Y., & Matei, A. (2016). *sampling: Survey Sampling*. R package.
CRAN.
:::
