---
title: "Detroit Transportation Survey: Multi-Stage Area Probability Sample Design"
subtitle: "A Technical Sampling Report"
author:
  - name: Effect Size
    affiliations: University of Maryland
    email: email.umd.edu
    corresponding: true
date: today
format:
  apaquarto-pdf:
    documentmode: man
    keep-tex: false
    floatsintext: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

### These are the packages we will be using. The [tidycensus](https://walker-data.com/tidycensus/) is the only package that requires an api key

```{r}
#| label: setup
#| echo: false
#| message: false
#| warning: false

# ──────────────────────────────────────────────────────
# Load Required Packages
# ──────────────────────────────────────────────────────
library(tidyverse)      # Data wrangling and visualization
library(readxl)         # Read Excel files
library(sampling)       # Survey sample selection (UPsystematic, etc.)
library(survey)         # Survey analysis
library(srvyr)          # Tidyverse-friendly survey analysis
library(PracTools)      # Practical tools for survey design
library(sf)             # Spatial data handling
library(tigris)         # Census TIGER shapefiles
library(tidycensus)     # Census data API
library(viridis)        # Color-blind friendly palettes
library(patchwork)      # Combine ggplots
library(knitr)          # Table formatting
library(kableExtra)     # Enhanced tables
library(scales)         # Number formatting


# tigris Options
options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

# set random seed
set.seed(-77)
```

#### Once you request your api key, you will need to save it locally and not share it in any of the scripts that will go on github. We may also use the stadia package for plots

```{r}

# paths may vary
#api_keys <- read_csv("D:/repos/api-keys.csv", show_col_types = FALSE)

```

### The data for this project can be read in through through github

```{r}

# data should now be read in automatically. 
input_path <- "C:/Users/tavri/Documents/SURV745/eight-mile-sample/inputs/"

census_tract <- "Census_Tract_FIPS_Code_Detroit_MI.csv"
detroit_mi <- "Detroit_MI.xlsx"


# read in census tracks
tract_fips <- read_csv(
  str_c(input_path, # change this as needed
        census_tract), col_names = "GEOID",
  col_types = cols(GEOID = col_character()),
  skip = 1
)

# read in detroit tracks
tract <- read_excel(
  str_c(input_path,
        detroit_mi), sheet = "Tract") |>
  mutate(Tract = as.character(Tract)) |>
  semi_join(tract_fips, by = c("Tract" = "GEOID"))

bg <- read_excel(
   str_c(input_path,
        detroit_mi), sheet = "BlockGroup") |>
  mutate(
    BlockGroup = as.character(BlockGroup),
    Tract = substr(BlockGroup, 1, 11)
  ) |>
  semi_join(tract_fips, by = c("Tract" = "GEOID"))

cat("Detroit tracts:", nrow(tract), "\n")
cat("Detroit block groups:", nrow(bg), "\n")
```

\newpage

# Introduction {#sec-intro}

This report will capture the the design of a multi-stage area probability sample for a face-to-face survey of Detroit, Michigan residents. The survey examines transportation security and transportation use, with key outcome variables including

-   the proportion of residents in households without a vehicle

-   the proportion whose main transportation is public transit, the proportion who skip destinations due to transportation problems

-   the proportion who sought transportation help from friends, family, or neighbors.

The sample uses a three-stage design:

1.  Stage 1: Census tracts as primary sampling units (PSUs),
2.  Stage 2: Block groups (BGs) as secondary sampling units (SSUs), and
3.  Stage 2: Persons as elements.

### Design Goals

The design goals are to yield 1,000 total respondents (see table below) while achieving a self-weighting sample within each age group and equal interviewer workload in each sample PSU. These dual objectives are accomplished through the composite measure of size (MOS) technique described in VDK.

### Deliverable Requiremetns

The report will be structured for a dual audience: managers seeking a high-level understanding of the sample design and its rationale, and survey statisticians requiring the mathematical details, selection probabilities, and variance estimation guidance.

# Sample Design {#sec-design}

## Design Goals

The sample must satisfy four requirements simultaneously:

1.  **Total sample size**: 1,000 respondents split equally (250 per group) across four age domains.
2.  **Self-weighting within domains**: Every person in a given age group has the same overall selection probability, i.e., $\pi_k(d) = f_d$ for domain $d$.
3.  **Equal workload per PSU**: The total expected number of selected persons is the same across all 40 sample tracts.
4.  **Multi-stage selection**: 40 tracts (PSUs), 1 BG per tract (SSU), and persons selected within BGs at domain-specific rates.

## Target Population

The target population consists of all non-institutionalized persons aged 18 and older residing in the City of Detroit, Michigan. According to the 2020 Decennial Census Demographic and Housing Characteristics (DHC) Summary File, Detroit contains approximately 480,131 adults (18+) across 270 eligible census tracts.

## Area Sampling Frame {#sec-frame}

### Data Sources

The sampling frame is constructed from two files:

-   **Census Tract FIPS Code Detroit MI.csv**: A list of 280 FIPS codes identifying Census tracts within the City of Detroit, obtained from the City of Detroit Open Data Portal.

-   **Detroit MI.xlsx**: Census tract- and block group-level aggregate data from the 2020 Decennial Census DHC Summary File, downloaded via the `tidycensus` R package for Wayne County, MI and merged with the Detroit tract FIPS codes. Includes housing unit counts (`H1_001N`), total person counts, and detailed person counts by Sex $\times$ Age (`P12` table).

### Data cleaning (still in development)

First, there appears to be a tract (26163985500) that might be a body of water and has 0 population and shows up identically 25 times in the tract dataset. We will drop it since it does not contribute to our analysis.

```{r}
tract |>
  filter(Tract == "26163985500") |> 
  select(1:10) |> 
  head() |> 
  kable()

# drop this tract
tract <- tract |> 
  filter(Tract != "26163985500")

tract |> nrow() |> kable()
```

-   We can now restructure the datasets by PSU and SSU.

```{r}

# bin age based on what was given to us 
age_patterns <- c(
  "Age18to34" = "18to19|20Yrs|21Yrs|22to24|25to29|30to34",
  "Age35to49" = "35to39|40to44|45to49",
  "Age50to64" = "50to54|55to59|60to61|62to64",
  "Age65plus" = "65to66|67to69|70to74|75to79|80to84|GE85"
)

# PSU Frame (tract-level)
## first, pivot longer by gender and age
psu_frame <- tract |>
  pivot_longer(
    cols = matches("^Male[^T]|^Female[^T]"),
    names_to = "SexAge",
    values_to = "elements"
  ) |>
  mutate(
    # using age_pattern, create recode age
    AgeGroup = map_chr(SexAge, \(x) {
    match <- names(age_patterns)[str_detect(x, age_patterns)]
    if (length(match) == 0) NA_character_ else match
  })
) |>
  filter(!is.na(AgeGroup)) |>
  # collapse elements to columns
  summarise(n = sum(elements), .by = c(Tract, TotHH, TotPerson, AgeGroup)) |>
  pivot_wider(names_from = AgeGroup, values_from = n)

# SSU Frame (block group-level)
ssu_frame <- bg |>
  pivot_longer(
    cols = matches("^Male[^T]|^Female[^T]"),
    names_to = "SexAge",
    values_to = "elements"
  ) |>
  mutate(
    AgeGroup = map_chr(SexAge, \(x) {
      match <- names(age_patterns)[str_detect(x, age_patterns)]
      if (length(match) == 0) NA_character_ else match
    })
  ) |>
  filter(!is.na(AgeGroup)) |>
  # collapse elements to columns
  summarise(n = sum(elements), .by = c(BlockGroup, Tract, TotHH, TotPerson, AgeGroup)) |>
  pivot_wider(names_from = AgeGroup, values_from = n)



# ── Check if this structure is correct for what we need

```
